---
title: "p8105_hw6_lz2949"
author: "Longyi Zhao"
date: "2023-11-25"
output: github_document
---
```{r}
library(tidyverse)
library(modelr)
library(p8105.datasets)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)

theme_set (theme_minimal() +theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplots.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_disrete = scale_fill_viridis_d

set.seed(1)
```


## Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


## Problem 2
```{r}
# data import 
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

```{r}
# fit the regression model 
tmax_fit = weather_df |>
  lm(tmax ~ tmin + prcp, data = _)
```

```{r}
# run a bootstrap
tmax_boot = weather_df |> 
  modelr::bootstrap(n = 5000) |> 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    results = map(models, broom::tidy),
    glance = map(models, broom::glance)) |>
  select(results, glance) |>
  unnest(results,glance)

r_squared_values = tmax_boot |>
  filter(term == "(Intercept)") |>
  select(r.squared)

log_beta_values = tmax_boot |>
  filter(term == "tmin"| term == "prcp") |>
  select(estimate) 

log_product_value = mutate(log_beta_values, group = rep(1:(n() / 2), each = 2)) |>
  group_by(group) |>
  summarize(log_product = log(prod(estimate)))

# MANY NaN
```

```{r}
# confidence interval r^2
r_interval = r_squared_values |>
  summarize(ci_lower = quantile(r.squared, 0.025), 
            ci_upper = quantile(r.squared, 0.975))

cat("The 95% confidence interval for r-square is (", r_interval$ci_lower, ",", r_interval$ci_upper, ")\n")

# confidence interval log(B*B)
log_interval = log_product_value |>
  summarize(ci_lower = quantile(log_product, 0.025, na.rm = TRUE), 
            ci_upper = quantile(log_product, 0.975, na.rm = TRUE))

cat("The 95% confidence interval for log products is (", log_interval$ci_lower, ",", log_interval$ci_upper, ")\n")
```

```{r}
# plot the distribution of estimates
estimated_data = cbind(r_squared_values, log_product_value) 

ggplot(estimated_data, aes(x = r.squared)) +
  geom_density() + 
  labs(title = "Distribution plot of r.squared")
```
Interpretation: r-squared typically ranges from 0 to 1, where 0 indicates the model does not explain any of the variance, and 1 indicates the model explains all of the variance. The plotted r squared distribution is close to normal, centered around 0.92, showing the model fits the data in a pretty good way. 

```{r}
ggplot(estimated_data, aes(x = log_product)) +geom_density() +
  labs(title = "Distribution plot of log product")
# report how many NaN in the dataset
```
Interpretation: Lots of Beta2 are negative values, leading to log(negative numbers) which does not exist in real number space. Among 5000 log products, `r estimated_data |> filter(log_product == "NaN") |> count()` of them are NaN. Which counts for `r (estimated_data |> filter(log_product == "NaN") |> count())/5000*100` percent of 5000 data points. The distribution plot of log products is clearly not normally distributed, it is left-skewed and has a heavy tail extending to low values, values around -6 has the highest density. Features like that may be related to the frequency with with large outliers are included in the bootstrap sample. 


## Problem 3
Analyze data gathered to understand the effects of several variables on a child's birthweight.
```{r}
# import data
bwt_df = read_csv("./data/birthweight.csv") |>
  janitor::clean_names() |>
  drop_na() |>
  mutate( 
    babysex = factor(
      babysex, levels = c("1","2"),labels = c("male", "female")), 
    frace = factor(frace, levels = c("1","2","3","4","8","9"), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")), 
    mrace = factor(mrace, levels = c("1","2","3","4","8"), labels = c("White", "Black", "Asian", "Puerto Rican", "Other")), 
    malform = factor(malform, levels = c("0","1"), labels = c("absent", "present")))
# drop NA
# convert sex/frace/mrace/malform to factor
```

Propose a regression model for birthweight. use add_predictions and add_residuals in making plots. 

```{r}
bwt_df |>
  ggplot(aes(x = wtgain, y = bwt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ylab("Birth weight of baby (grams)") +
  xlab("Mother's weight gain during pregnancy (pounds)")+
  labs(title = "plot of mother's weight gains vs. baby birth weight")
```

```{r}
bwt_df |>
  ggplot(aes(x = mheight, y = bwt)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  ylab("Birth weight of baby (grams)") +
  xlab("Mother's height (inches)")+
  labs(title = "plot of mothers' height vs. baby birth weight")
```

```{r}
# linear model 
prop_lm = lm(bwt ~ wtgain + mheight, data = bwt_df)


bwt_df |>
  modelr::add_residuals(prop_lm) |>
  modelr::add_predictions(prop_lm) |>
  ggplot(aes(x = pred, y = resid)) + geom_point() +
  ylab("Residual") +xlab("Fitted values") +labs(title = "plot of residual against fitted values")
```
Description: Ideas that mothers who experience a more substantial weight gain during pregnancy are likely to deliver babies with higher birth weights and higher mothers probably have greater capacities of bearing bigger babies in them immediately come to my mind. I first plot the baby birthweight vs. mother's weight gain during pregnancy and fit a line to see the trend, the line has a positive slope, indicating that there might be a positive correlation between mother's weight gain and birth weight of baby. Then I did similar stuff to the mother's height. Finally I build the slr using mothers' weight gain and mothers' height as predictors. 
<br>
From the residual plot, it can be found that most points are gathered around 0 with no obvious trend or distribution. 

```{r}
# using length at birth and gestational age as predictor
lbga_mod = lm(bwt ~ blength + gaweeks, data = bwt_df)
# use head circumference, length, sex, and all interactions between these
hcls_mod = lm(bwt ~ bhead + blength + babysex + bhead * blength * babysex, data = bwt_df)
```

```{r}
# comparison in terms of cross-validated prediction error; use corssv_mc and functions in purrr. 
b_df =
  crossv_mc(bwt_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble))

# use mutate +map and map 2 to fit models
b_df = 
  b_df |> 
  mutate(
    prop_lm  = map(train, \(df) lm(bwt ~ wtgain + mheight, data = df)),
    lbga_mod = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    hcls_mod = map(train, \(df) lm(bwt ~ bhead + blength + babysex + bhead * blength * babysex, data = df))) |> 
  mutate(
    rmse_my_model = map2_dbl(prop_lm, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_birthlength_gestationalage = map2_dbl(lbga_mod, test, \(mod, df) rmse(model = mod, data = df)),
    rmse_headcirc_length_sex = map2_dbl(hcls_mod, test, \(mod, df) rmse(model = mod, data = df)))
```

```{r}
# plot the preiction error 
b_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  xlab("Models") + ylab("rmse values") +labs(title = "rmse values' Violin plots for three models")
```

Interpretation: RMSE is a measure of the average deviation between the observed values and predicted values by the regression model. A lower RMSE indicates a better fit of the model to the data. From the violin plot, it can be see that the second and third models are better than the first one, they have lower RMSE values, indicating they have smaller prediction error. This is probably because the first model use mothers' characteristics as the predictor which does not directly impact baby's birth weight; the second and third use babies' characteristics as predictors when building the model which have a more direct correlation and can explain more variances of the outcome and thus fit data in a better way. 





